---
title: "503_Nordic_database_sites"
author: "DHJ"
date: "6 3 2020"
output: html_document
---

## 1. Libraries
```{r}

library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(readxl)
library(lubridate)

```

## 2. Data
```{r}

# Chemical data (which also are the stations of befgroing and bunndyr)
dat <- read_excel("Nordisk_database_data/Resultater 2019.xlsx", sheet = "Stasjoner")

# We use this for Vannforekomst ID (Surface Water Body Code)
df_typifisering <- read_excel("Nordisk_database_data/Typifiseringsparametere 2019 snitt, sd osv.xlsx", 
                      sheet = "StationAttribute")

# Vannmiljø data
# All data in project "Referanseelver"  
df_vm <- read_excel("Nordisk_database_data/Vannmiljo_2017_2019.xlsx", guess_max = 38000)

# NIVAbase (from script 505)
df_nivabase <- read.csv("Nordisk_database_data/Stations - 2017-18.csv")

# Water body info (info on "National type") 
df_waterbodies <- read_excel("K:/Prosjekter/Vanndirektivet/Nordisk fys-kjem QE/SOW_SWB_SurfaceWaterBody_NO.xlsx")

# River/catchment data
df_riverinfo <- read_excel("Nordisk_database_data/Stations and catchment data all rivers.xlsx")


```

### Check Vannmiljø  
Conclusion is that fish belong to separate stations, while both chemistry, påvekstalger and   
bunndyr belong to a single station
```{r}

if (FALSE){
  
  # Shows types of medium for each station 
  df_vm %>%
    group_by(Vannlokalitet_kode, Vannlokalitet) %>%
    summarise(n = n(), Medium = paste(unique(Medium_navn), collapse = ",")) %>% View()
  
  df_vm %>%
    count(Parameter_navn)
  
  # Check some sample parameters  
  # And see which water locality they belong to
  df_vm %>% 
    filter(Parameter_navn %in% c("Begroing EQR eutrofiering", "Fosfat (ufiltrert)",
                                 "Average Score per Taxon (ASPT)", "Estimert totalt antall fisk per arealenhet")) %>%
    arrange(Parameter_navn) %>%
    group_by(Vannlokalitet_kode, Vannlokalitet) %>%
    summarise(n = n(), Medium = paste(unique(Parameter_navn), collapse = ",")) %>% View()
  
}


```

## 3. Add Vannmiljø columns to Nivabase stations  
Not an obvious way to connect Vannmiljø and NIVA stations as both codes and names differ.  


### Add UTM coordinates to df_nivabase  
```{r}

# Vannmiljø sites
df_vm_sites <- df_vm %>%
  count(Vannlokalitet_kode, Vannlokalitet, `UTM33 Ost (X)`, `UTM33 Nord (Y)`)

#
# Add UTM coordinates to df_nivabase
#
library(sp)
crs_longlat <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
crs_utm <- "+proj=utm +zone=33 +ellps=WGS84 +datum=WGS84 +units=m"
 
SP <- SpatialPoints(df_nivabase[,c("LONGITUDE", "LATITUDE")],
         proj4string=CRS(crs_longlat)
         )
SP.UTM <- spTransform(SP, CRS(crs_utm))
# Add transformed coords to data set
df_nivabase$`UTM33 Ost (X)` <- SP.UTM@coords[,1]
df_nivabase$`UTM33 Nord (Y)` <- SP.UTM@coords[,2]

```

### Test plot - NIVAbase + Vannmiljø points
```{r}

ggplot(df_nivabase, aes(x = `UTM33 Ost (X)`, y = `UTM33 Nord (Y)`)) +
  geom_point() +
  geom_point(data = df_vm_sites, color = "blue",size = 3, pch = 1)

```

### Find closest Vannmiljø station     
For each station in 'df_nivabase', find closest stations in Vannmiljø  
```{r}

#
# Function: Find which observation in dat2 is closest to observation'i' in dat1
# Returns index (of dat2 observation) and distance (if UTM, in meters)
#
find_closest <- function(i, dat1, dat2, x = "UTM33 Ost (X)", y = "UTM33 Nord (Y)"){
  dist <- sqrt((dat1[[x]][i] - dat2[[x]])^2 + (dat1[[y]][i] - dat2[[y]])^2)
  data.frame(Index = which.min(dist), Dist = min(dist) %>% round(1))
}
# Test
# find_closest(1, df_nivabase, df_vm_sites)

# For each station in 'df_nivabase', find closest stations in Vannmiljø  
# Returns data frame as long as 'df_nivabase', with Index telling which 
#   df_vm_sites row to pick
df_closest <- df_nivabase %>% 
  nrow() %>%
  seq_len() %>%
  map_df(find_closest, dat1 = df_nivabase, dat2 = df_vm_sites)

# Pick rows of df_vm_sites row, and the two relevant columns  
columns_to_add <- df_vm_sites[df_closest$Index, c("Vannlokalitet_kode", "Vannlokalitet")]

# Add columns Index and Dist
columns_to_add <- columns_to_add %>% bind_cols(df_closest)
# columns_to_add

# Remove (set to NA) Vannmiljø stations >1 km away form NIVA stations
sel <- columns_to_add$Dist > 1000
columns_to_add[sel,] <- rep(NA, 4)

# Show result
columns_to_add

```

### Add Vannmiljø columns to df_nivabase    
Added by just adding the columns   
```{r}
# Added by just add in the columns   
df_nivabase <- df_nivabase %>% bind_cols(columns_to_add)

# For check
if (FALSE){
  df_nivabase %>%
    select(STATION_ID, STATION_NAME, Vannlokalitet_kode, Vannlokalitet) %>% View()
}

# Save
openxlsx::write.xlsx(df_nivabase, "Nordisk_database_data/NIVA and corresponding Vannmiljø stations.xlsx")

```

## 4. Add columns to dat


### Add VannforekomstID   
From df_typifisering  
```{r}

dat <- dat %>% 
  left_join(
    df_typifisering %>% select(StationId, VannforekomstID), # , Vannregion, `Vannregionens EU-CD`),
    by = "StationId"
  )

# sel <- is.na(dat$VannforekomstID)
# sum(sel)
# dat[sel,]

# Some VannforekomstID must be adjusted  
dat <- dat %>% 
  mutate(VannforekomstID = case_when(
    StationCode %in% "O_002-305_Atn11" ~ "002-305-R",
    StationCode %in% "O_002-305_Atn4"  ~ "002-305-R",
    StationCode %in% "O_002-2572_Lun"  ~ "002-2572-R",
    StationCode %in% "O_002-604_Vik"   ~ "002-604-R",
    TRUE ~ VannforekomstID
  ))

sel <- is.na(dat$VannforekomstID)
sum(sel)

```

### Add European VannforekomstID   
Just add "NO" in front of it  
```{r}

dat <- dat %>%
  mutate(VannforekomstID_EU = paste0("NO", VannforekomstID))

```


### Add Vannlokalitet_kode from Vannmiljø   
Will become "National ID for lake/river"  
```{r}

dat <- dat %>% 
  left_join(
    df_nivabase %>% select(STATION_ID, Vannlokalitet_kode, Vannlokalitet), # , Vannregion, `Vannregionens EU-CD`),
    by = c("StationId" = "STATION_ID")
  )

```



### Add Vannregion/vannområde
```{r}

dat_vann_nett <- read_excel("Nordisk_database_data/Vann-nett elv.xlsx", guess_max = 19000)
# write.table(dat_vann_nett, "Nordisk_database_data/Vann-nett elv.tsv")

```

### Check if all VannforekomstID are in data_vann_nett
```{r}

x <- dat$VannforekomstID 
sel <- x %in% dat_vann_nett$VannforekomstID

sum(!sel)  # should be zero
# x[!sel]

```

### Add Vannregion + vannområde   
```{r}

dat <- dat %>% 
  left_join(dat_vann_nett %>% select(VannforekomstID, 
                                     VannregionID, Vannregion,
                                     VannområdeID, Vannområde),
            by = "VannforekomstID")

# dat %>%  View()


```

## 5. Add 'National Type'  

Get data from the given excel file (tip from Kari)   
```{r}

df_waterbodies_sel <- df_waterbodies %>%
  filter(euSurfaceWaterBodyCode %in% dat$VannforekomstID_EU)

sel <- dat$VannforekomstID_EU %in% df_waterbodies_sel$euSurfaceWaterBodyCode
sum(!sel)  # should be zero

```

### Add to data
```{r}

dat <- dat %>%
  left_join(df_waterbodies_sel %>% select(euSurfaceWaterBodyCode, surfaceWaterBodyTypeCode),
            by = c("VannforekomstID_EU" = "euSurfaceWaterBodyCode"))

```



## 6. Add non-mandatory river/catchment info  

### Check 
```{r}

# Fix df_riverinfo$Kode
df_riverinfo <- df_riverinfo %>% 
  mutate(Kode = case_when(
    Kode %in% "O_002-305_Atn_DAN11" ~ "O_002-305_Atn11",
    Kode %in% "O_002-305_Atn_DAN04" ~ "O_002-305_Atn4",
    TRUE ~ Kode
  ))

sel <- dat$StationCode %in% df_riverinfo$Kode
sum(!sel) # should be 0

# If not all are found - fix df_riverinfo$Kode as shown above
# dat$StationCode[!sel]
# df_riverinfo$Kode %>% sort()

```


### Add columns from 'df_riverinfo'  
```{r}

dat <- dat %>%
  left_join(df_riverinfo %>% select(Kode, HEIGHT_50, AREAL_KM2, `AVRENNING (mm/år)`),
            by = c("StationCode" = "Kode"))

```

## 7. Put together mandatory columns   
we include only the following non-mandatory columns:
- "Site name" (we use Vannmiljø name here, not NIVA name)  
- Altitude (median of river)  
- Catchment size
- Runoff  
```{r}

dat_formatted <- dat %>% 
  mutate(Category = "River", 
         Country = "Norway",
         `EUID for site` = paste0("NO", Vannlokalitet_kode),
         `Monitoring program` = "Referanseelver",
         `IC Type` = "",
         `Lake size` = NA, `Lake mean depth` = NA, `Lake max depth` = NA, 
         `Clay impact` = NA, `Point sources of nutrients` = NA, `Water level fluctuaion` = NA, `Connectivity` = NA,
         `Ecological status` = as.character(NA), `Impact Acidification` = NA, 
         `Impact Eutrophication`  = NA, `Comments` = as.character(NA)
         ) %>%
  rename(`Unique Site ID`= StationId,
         `National Type` = surfaceWaterBodyTypeCode,
         `National ID for lake/river` = VannforekomstID,
         `EUID for water body` = VannforekomstID_EU,
         `Sub unit code` = VannområdeID, 
         `Sub unit name` = Vannområde,
         `River basin district code` = VannregionID, 
         `River basin district name` = Vannregion,
         `Site name` = Vannlokalitet,          # we use Vannmiljø name here, not NIVA name
         Altitude = HEIGHT_50,
         `Catchment size` = AREAL_KM2,
         Runoff = `AVRENNING (mm/år)`
) %>%
  # Order columns correctly
  select(`Unique Site ID`, Latitude, Longitude, Category,
         `National Type`, Country, `EUID for site`, `EUID for water body`, `National ID for lake/river`, 
         `Sub unit code`, `Sub unit name`, `River basin district code`, `River basin district name`,
         `Site name`, `IC Type`, `Monitoring program`, Altitude, `Catchment size`, 
         `Lake size`, `Lake mean depth`, `Lake max depth`, `Runoff`,
         `Clay impact`, `Point sources of nutrients`, `Water level fluctuaion`, `Connectivity`,
         `Ecological status`, `Impact Acidification`, `Impact Eutrophication`, `Comments`) 

View(dat_formatted)

#
# MANDATORY
#
# Unique Site ID
# Latitude DecDeg
# Longitude DecDeg
# Category (lake/river)
# National Type
# Country
# EUID for  site
# EUID for water body
# National ID for lake/river
# Sub unit code
# Sub unit name
# River basin district code
# River basin district name
#
# NON-MANDATORY
#
# Site name		Local name of site
# IC Type		According to intercalibration work
# Monitoring program. 		Name or type of monitoring program
# Altitude 	m s a l	
# Catchment size 	km2	
# Lake size	km2	
# Lake mean depth	m	
# Lake max depth (estimate/model)	m	
# Runoff	m/yr	30 year mean as late as possible).
# Clay impact	y/n	Please write how this was  defined.
# Point sources of nutrients 	y/n	
# Water level fluctuaion	y/n	
# Connectivity	y/n	y means that conectivity is impacted.
# Ecological status 		As reported to commission
# Impact Acidification	y/n	According to  impact assessment
# Impact Eutrophication	y/n	According to  impact assessment
# Comments	Free text	Any information in the site that might be useful
#



```



## 8. Save  
```{r}

openxlsx::write.xlsx(dat_formatted, "Nordisk_database_formatted/NIVA reference rivers sites 2017-2019.xlsx")

```


## 9. Appendix. Check shapefile of European Vannforekomster  
```{r}

library(sf)
# Norway (coastline) in UTM coordinates
shape_lake <- st_read("WISE WFD shapefiles/eea_v_4326_250_k_wise-wfd2016_p_2016-2019_v01_r03/shapefile/SurfaceWaterBody/SurfaceWaterBody.shp")

shape_river <- st_read("WISE WFD shapefiles/eea_v_4326_250_k_wise-wfd2016_p_2016-2019_v01_r03/shapefile/SurfaceWaterBody/SurfaceWaterBodyLine.shp")

# st_crs(shape) <- "+proj=utm +zone=33"  # set coordinate system


```

### Check some
```{r}

# 4 random lakes
shape_lake %>%
  filter(nameTxtLan == "nor") %>% head(4)

# All rivers including Glomma 
shape_river %>%
  filter(grepl("GLOMMA", nameText))

```


4