---
title: "504_Nordic_database_begroing_indices"
author: "DHJ"
date: "6 3 2020"
output: html_document
---

Indices for begroing 
- Uses data from script 505 for 2017-2018 data  

## 1. Libraries
```{r}

library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(readxl)
library(lubridate)

source("501_Nordic_database_chemistry_functions.R", encoding = "UTF-8")

```


## 2. 2019 data  
```{r}

dat_2019 <- read_excel("Nordisk_database_data/Begroingsalger PIT og AIP - 2019.xlsx") %>%
  mutate(Indeksverdi = as.numeric(Indeksverdi), 
         EQR = as.numeric(EQR),
         nEQR = as.numeric(nEQR))

# Data from Access database  
dat_2019_raw <- read_excel("Nordisk_database_data/Begroing RefElver data 2019.xlsx", sheet = "dates and site codes")
dat_2019_codes <- read_excel("Nordisk_database_data/Begroing RefElver data 2019.xlsx", sheet = "site names and codes", skip = 2)
  
```

### Put on long format  
```{r}

dat_ind <- dat_2019 %>% 
  select(-c(EQR, nEQR)) %>% 
  rename(Value = Indeksverdi)
dat_EQR <- dat_2019 %>% 
  select(-c(Indeksverdi, nEQR)) %>%
  rename(Value = EQR) %>% 
  mutate(Indeks = paste(Indeks, "EQR"))
dat_nEQR <- dat_2019 %>% 
  select(-c(Indeksverdi, EQR)) %>%
  rename(Value = nEQR) %>% 
  mutate(Indeks = paste(Indeks, "nEQR"))

dat_2019 <- bind_rows(dat_ind, dat_EQR, dat_nEQR)

```

### Get data for STATION_ID  
```{r}

#
# Fix stations 
#
sel <- dat_2019$`Aquamonitor St. kode` %in% "O_002-305_Atn4"; sum(sel)
dat_2019$`Aquamonitor St. kode`[sel] <- "O_002-305_Atn_DAN04" 

sel <- dat_2019$`Aquamonitor St. kode` %in% "O_002-305_Atn11"; sum(sel)
dat_2019$`Aquamonitor St. kode`[sel] <- "O_002-305_Atn_DAN11"

#
# Station data
#
df_stations <- read.csv("Nordisk_database_data/Stations - 2017-18.csv")

#
# Check if all codes in dat_2019 are present in df_stations
#
codes <- unique(dat_2019$`Aquamonitor St. kode`)
sel <- codes %in% df_stations$STATION_CODE
sum(!sel)  # should be zero 

# IF NOT all codes are present, fix them in dat_2019 (see above) 
# codes[!sel]
# df_stations$STATION_CODE %>% sort()

```

### Add STATION_ID  
```{r}

dat_2019 <- dat_2019 %>%
  left_join(df_stations %>% select(STATION_CODE, STATION_ID), 
            by = c("Aquamonitor St. kode" = "STATION_CODE"))


```

### Add dates  
- Must first find Access code for the site ('lok_sta') from dat_2019_codes  
- Can then get date from dat_2019_raw

#### Check codes
```{r}

# Fix one
dat_2019_codes <- dat_2019_codes %>%
  mutate(st_kode = case_when(
    st_navn %in% "Øydgardselva i Volda, RefElver" ~ "V_094-102_Øyd",
    TRUE ~ st_kode
  ))


codes <- dat_2019$`Aquamonitor St. kode` %>% unique()
sel <- codes %in% dat_2019_codes$st_kode
sum(!sel) # should be zero

# if not zero:
if (FALSE){
  codes[!sel]
  dat_2019_codes$st_kode %>% sort()
  
  # For copy example to excel
  dat_2019 %>%
    filter(`Aquamonitor St. kode` %in% codes[!sel] &  Indeks == "PIT") %>%
    write.table("clipboard", sep = "\t", dec = ",", row.names = FALSE)
}

# dat_2019_codes$lok_sta %>% sort()
# dat_2019_raw$lok_sta %>% unique() %>% sort()

# "O_002-305_Atn_DAN04" "O_002-300_Atn_DAN03" "O_002-305_Atn_DAN11"


```

#### Add Access database code  
```{r}

dat_2019 <- dat_2019 %>%
  left_join(dat_2019_codes %>% select(st_kode, lok_sta),
            by = c("Aquamonitor St. kode" = "st_kode"))


```

#### Add date
```{r}
dat_2019_dates <- dat_2019_raw %>%
  count(lok_sta, dato)

# Check if there is only one date per station 
dat_2019_dates %>%
  count(lok_sta) %>%
  filter(n > 1) %>%
  nrow()     # should be zero

# Add date
dat_2019 <- dat_2019 %>%
  left_join(dat_2019_dates %>% select(lok_sta, dato))

```

### Fix remaining dates    
Date gotten on mail from Susi  
```{r}

# dat_2019 %>% filter(is.na(dato)) %>%
#   count(`Aquamonitor St. kode`)

sel <- dat_2019$`Aquamonitor St. kode` %in% c("O_002-1869_Dor", "O_002-300_Atn_DAN03", 
                                              "O_002-305_Atn_DAN04", "O_002-305_Atn_DAN11")
sum(sel)
dat_2019$dato[sel] <- "2019-08-30 UTC"


```


### Put on correct format  
```{r}

data_2019_formatted <- dat_2019 %>%
  mutate(Year = year(dato),
         Month = month(dato),
         Day = day(dato)) %>%
  rename(`Unique site ID` = STATION_ID,
         Index = Indeks) %>%
  select(`Unique site ID`, Year, Month, Day, Index, Value)

```



## 3. 2017-18 data
```{r}

data_2017_18 <- read.csv("Nordisk_database_data/Begroingsalger PIT og AIP - 2017-18.csv")

data_2017_18_formatted <- data_2017_18 %>%
  mutate(Year = year(SAMPLE_DATE),
         Month = month(SAMPLE_DATE),
         Day = day(SAMPLE_DATE)) %>%
  rename(`Unique site ID` = STATION_ID,
         Index = NAME,
         Value = VALUE) %>%
  select(`Unique site ID`, Year, Month, Day, Index, Value)

```

## 4. Data formatted  
```{r}

data_formatted <- bind_rows(data_2017_18_formatted, data_2019_formatted)

xtabs(~Year + Index, data_formatted)

```

## 5. Save  
```{r}

openxlsx::write.xlsx(data_formatted, "Nordisk_database_formatted/NIVA reference rivers BD diatoms.xlsx")

```

